---
title: "Statistical Inference Course Project: The Central Limit Theorem (CLT) and Simulation Experiment"
author: "Michela Ieva"
date: "Thursday, May 21, 2015"
output: pdf_document
---

## Overview
We report about the results of a CLT simulation experiment. In this analyasis we will investigate the distribution of averages of 40 exponentials, performing 1000 simulations.
According the CLT the resulting distribution looks like a bell curve with mean and standard deviation compatible with the theoretical values of *Normal* distribution.

```{r setup}
library(knitr)
library(ggplot2)

echo = TRUE  # Always make code visible

```

## Simulation
The exponential distribution can be simulated in R with the function rexp(n, lambda) where lambda is the rate parameter. The theoretical mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. 

First of all, for reproducibility we need to seet the seed than we set the simulation parameters.

```{r}
set.seed(1234) # for reproducibility purpose

nosim <- 1000
n <- 40
lambda = 0.2
meanexp = 1/lambda
sigmaexp = 1/lambda

```

We use the funcion *replicate* to perform a thousand of simulations of a sample of 40 exponential distributions and take the mean of each sample.

```{r}
dat <- replicate(nosim, mean(rexp(n, lambda)))

```

Our dataset looks like:

```{r}
str(dat, 10)

```

## Data analysis
In order to study the behaviour of the averages distribution we plot the simulated dataset.

```{r plot1, fig.height= 4, fig.width= 4}

h = hist(dat, prob=TRUE, main = "Histogram for simulated means", xlab = "Means of each sample")
curve(dnorm(x, mean=mean(dat), sd=sd(dat)), add=TRUE, col = "green")

```

As you can see the distribution looks like a bell curve. We superimpose, futhermore a *normal* distribution (green line) with the mean and standard deviation coming from our simulated dataset.

### Sample Mean versus Theoretical Mean
Using the simulated sample the mean is:

```{r sample variables}
# sample mean
mean_dat <- mean(dat)
mean_dat

```
Instead, as known, the theoretical mean of the distribution is supposed to be `1/lambda`:

```{r population mean}
mean_theor<-1/lambda
mean_theor

```

The two values agree very well. Furthermore we can compare theoretical and simulated parameters using the 95% confidence interval for the averages.

```{r CI}
mean_dat + c(-1, 1)*qnorm(0.975)*sd_dat/sqrt(n)

```
As you can see the theoretical value lies perfectly inside this interval.

### Sample Variance versus Theoretical Variance

```{r theoretical variables}
# sample standard deviation and variance 
sd_dat <- sd(dat)
sd_dat

var_dat <- var(dat)
var_dat


sd_theor <- (1/lambda)/sqrt(40)
sd_theor

var_theor <- sd_theor^2
var_theor

```

Also in this case sample and theorethical values agree very well.

### Comparison with Guassian distribution

In order to compare theoretical and simulated distribution we superimpose to the previous plot a *normal* distribution with theoretical parameters (red line).

```{r plot2, fig.height= 4, fig.width= 4}
h = hist(dat, prob=TRUE, main = "Histogram for simulated means vs theoreticals", xlab = "Means of each sample")
curve(dnorm(x, mean=mean(dat), sd=sd(dat)), add=TRUE, col = "green")
curve(dnorm(x, mean=mean_theor, sd=sd_theor), add=TRUE, col = "red")

```

As you can see, the two distributions agree very well confirming the distribution is approximately normal.

## Conclusions
We studied the distribution of the averages of 1000 samples of 40 exponentials, by simulation. we found this distribution looks like a bell curve with sample parameters, mean and variance, agree very well with theorethical values.




